---
title: Long-term nutrient enrichment modifies wetland plant-microbe relationships

author: "Daniya Stephens, Regina Bledsoe, Ariane L. Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: null
header-includes:
- \usepackage{array}
- \usepackage{graphics}
- \usepackage[utf8]{inputenc}
---

Project Description: Fill out

# Initial Setup
```{r Initial Setup, include=FALSE}
rm(list=ls())
#PC - set WD manually by Session -> Set Working Directory -> Choose Directory...
if(Sys.info()[1] == "Darwin"){
  setwd("~/GitHub/WRC18_RhizoTeaDecomp/analyses")
} else {
  setwd(choose.dir())
}
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan")
require("dplyr")
require("tidyverse")
require("phyloseq")
require("agricolae")
require("ape")
require("picante")
require("reshape2")
require("reshape")
require("ade4")
require("labdsv")
require("ggpubr")
```
#Import Files
```{r Import Files - Env, include=FALSE}
#Load data files 
#Assign file names to variables
sharedfile = "../data/WRC18_Rhizo/WRC15_18Rhizo.opti_mcc.shared"
taxfile = "../data/WRC18_Rhizo/WRC15_18Rhizo.opti_mcc.0.03.cons.taxonomy"
metafile = "../data/WRC18_Rhizo/2015_2018_WestResearchCampus_Sampling_Summary.csv"
fungalfile = "../data/WRC18_Rhizo/fungal_otu_table_rhizo.csv"

#Read in design file
meta <- read.csv(metafile)
meta <- meta %>%
  filter(year=='2018')%>%
  filter(source=='bulk')%>%
  filter(treatment=='M'|treatment=='MF')%>%
  filter(block=='2'|block=='4'|block=='6'|block=='8')
```

```{r read in fungal OTUs}
#Read in fungal file
fungal_otu <- read.csv(fungalfile)

fungal_otu <- fungal_otu %>%
  column_to_rownames("OTU_ID")

fungal_otu <- t(fungal_otu)

#Removing mock community
fungal_otu <- fungal_otu[-c(1:2),]

#fungal taxa file - read in a clean up underscores (_) and "unidentified"
fungal_taxa <- read.delim("../data/WRC18_Rhizo/otu_taxa_table_v1.txt", header=T, sep="\t")

fungal_taxa_trim <- fungal_taxa %>%
  mutate(k = substr(kingdom, 4, str_length(kingdom)), p = substr(phylum, 4, str_length(phylum)),c = substr(class, 4, str_length(class)), o = substr(order, 4, str_length(order)),  f = substr(family, 4, str_length(family)), g = substr(genus, 4, str_length(genus)), s = substr(species, 4, str_length(species))) %>%
  dplyr::select(OTU_ID,domain=k,phylum=p,class=c,order=o,family=f,genus=g,species=s)%>%
  mutate(phylum = str_replace_all(phylum,"_", " "), class = str_replace_all(class,"_", " "), order = str_replace_all(order,"_", " "), family = str_replace_all(family,"_", " "), genus = str_replace_all(genus,"_", " "), species = str_replace_all(species,"_", " "))%>%
  mutate(phylum = str_replace(phylum,"unidentified", " "), class = str_replace(class,"unclassified", " "), order = str_replace(order,"unidentified", " "), family = str_replace(family,"unidentified", " "), genus = str_replace(genus,"unidentified", " "), species = str_replace(species,"unidentified", " "))
```

```{r Fungal Relative abundance and PERMANOVA}
b <- ncol(fungal_otu)
reads <- sum(colSums(fungal_otu))
#total reads before filtering = 229296

fungal_otu <- fungal_otu[, which(colSums(fungal_otu) > 10)]

fun_otu<- colSums(t(fungal_otu))
fun_otu
fun_otu[which.min(colSums(t(fungal_otu)))]
fun_otu[which.max(colSums(t(fungal_otu)))]

#Rarefy
#Set min sample number
min.N <- min(rowSums(fungal_otu))
#rrarefy returns a randomly rarefied community data frame or vector of selected size
#This is the otu set that should be used for beta diversity metrics
fungal_otu <- rrarefy(fungal_otu, min.N)

#subset bulk soil only
fungal_otu <- subset(fungal_otu[c(1:8),])

#transforms read counts to relative abundance
#Using rarefied OTU dataset
fungal_otu.rel <- fungal_otu
for(i in 1:dim(fungal_otu)[1]){
  fungal_otu.rel[i,] <- fungal_otu[i,]/sum(fungal_otu[i,])
}

#Combine experimental design meta data with OTU relative abundance data
fun_metaotu <- cbind(meta, fungal_otu.rel)
rownames(fun_metaotu) <- meta$sample
colnames(fun_metaotu)

#separate M vs MF for Mantel tests later
fun.2018.metaotu.M <- subset(fun_metaotu, treatment == "M")
fun.2018.metaotu.MF <- subset(fun_metaotu, treatment == "MF")

#remove metadata for fungal OTU by treatment only - use for Mantel test later
fun.2018.otu <- subset(fun_metaotu[,-c(1:26)])
fun.2018.otu.M <- subset(fun.2018.metaotu.M[,-c(1:26)])
fun.2018.otu.MF <- subset(fun.2018.metaotu.MF[,-c(1:26)])

#PERMANOVA with source and treatment as factors
#Among all treatments source (soil source) and treatment
fun_metaotu.ad <- adonis(fun_metaotu[,-c(1:26)] ~treatment, method = "bray", data = fun_metaotu, perm=1000, set.seed=42)

fun_metaotu.ad 
```

```{r Fungal Calculate distance matrix and PCoA components - Taxa/Otus}
fungal_otu.rel.dist<-vegdist(fungal_otu.rel, method="bray")

# Principal Coordinates Analysis

pcoa <- cmdscale(fungal_otu.rel.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #34.7
explainvar2b #18.8
```
```{r Ordination (PCoA) - fungal, include=FALSE}
# Principal Coordinates Analysis - even blocks only, mowed only, bulk only

pcoa.groups <- paste(fun_metaotu$treatment)
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt <- c("M","MF")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=trt)) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  scale_colour_manual(labels = c("mowed/unfertilized","mowed/fertilized"), 
                      values = c("gray70", "darkgreen")) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (34.7%)") + ylab("PCoA 2 (18.8%)") +
        labs(colour = "Treatment") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4))) 
    #+ ggtitle("Fungal Community Composition (2018)")
ggsave("../figures/2018.fung.ordination.bulk.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE) #do not change this code
```

```{r fungal Presence_Absence and Species richness, include =FALSE }
#Transform data to reflect only presence or absence ie 0 or 1
fun.otu.2.PA<-(fungal_otu.rel>0*1)
#sanity check 
#otu.2.PA[,1:5]
#Add meta data
fun.metaotu.2 <- cbind(meta, fun.otu.2.PA)
#metaotu.2[1:15]
#Transform True/False to 0/1 
cols <- sapply(fun.metaotu.2, is.logical)
fun.metaotu.2[,cols]<-lapply(fun.metaotu.2[,cols],as.numeric)
#head(metaotu.2)
```

```{r fungal diversity metrics: richness}
#chao1 richness index
fun.chao1 <- estimateR(fungal_otu)
c <- as.data.frame(fun.chao1)
ct <- as.data.frame(t(c))

meta$fun.s.obs <- ct$S.obs
meta$fun.s.chao1 <- ct$S.chao1
meta$fun.se.chao1 <- ct$se.chao1
colnames(meta) #total columns is 29

#interaction variable so can use Tukey
#i <- with(meta, interaction(treatment, source))

#plot fun.s.obs (OTU richness)
fun.richness.lm <- aov(fun.s.obs ~ treatment, data = meta)
plot(fun.richness.lm)
anova(fun.richness.lm)

#Plot
p <- ggplot(meta, aes(x=treatment, y=fun.s.obs, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
f_chao<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Fungal OTU richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized"))
    #+ ggtitle("Fungal OTU Richness (2018)")
f_chao
ggsave("../figures/2018.fungal.richness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE) #do not change this code
```

```{r diversity metrics - calculate fungal diversity and evenness}
# Calculate diversity metrics
fun.shannon <- vegan::diversity(fungal_otu, "shannon")
fun.simp <- vegan::diversity(fungal_otu, "simpson")
fun.invsimp <-vegan::diversity(fungal_otu, "invsimpson")
# Pielous evenness (uses presence/absence data)
fun.J <- fun.shannon/log(specnumber(fun.otu.2.PA[,-c(1:1)]))

#Adding diversity fields to dataset
meta$fun.shannon <- fun.shannon
meta$fun.J <- fun.simp
meta$fun.simp <- fun.invsimp
meta$fun.invsimp <- fun.J
colnames(meta) #total columns is 33
```

```{r diversity metrics - display fungal shannon }
#Run richness first. interaction, levels, and treatments for plotting set there
fun.rich.lm <- aov(fun.s.obs ~ treatment, data = meta)
plot(fun.rich.lm)
anova(fun.rich.lm)

fun.shannon.lm <- aov(fun.shannon ~ treatment, data = meta)
plot(fun.shannon.lm)
anova(fun.shannon.lm)

#Plot
p <- ggplot(meta, aes(x=treatment, y=fun.shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
f_shannon<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized")) + ggtitle("Fungal Diversity (2018)")
f_shannon
ggsave("../figures/2018.fungal.shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## Plant Data - 2018 only
```{r Import Files - Plant, include=FALSE}
PCC <- read.csv("../data/WRC18_plant_RA.csv", header=TRUE)
design <- read.csv("../data/2018.diversity.plant.csv", header=TRUE)
PCC.2018 <- subset(PCC, treatment == "M" | treatment == "MF")
PCC.2018 <- cbind(design$block,PCC.2018)
PCC.2018 <- PCC.2018 %>% filter(design$block=='2'|design$block=='4'|design$block=='6'|design$block=='8')

PCC.2018b <- PCC.2018[,-c(1:3)]
labels(PCC.2018b)
PCC.2018a <- PCC.2018b[, which(colSums(PCC.2018b) > 0)]
design.even <- PCC.2018[,c(1:3)]
PCC.2018.even <- cbind(design.even,PCC.2018a) 
```
# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Plant, echo=TRUE}
# Fisher's Alpha
PCC.2018.PA <- decostand(PCC.2018a, method="pa", na.rm=TRUE)

# Species Richness
richness <- rowSums((PCC.2018.PA >= 1))

# Shannon Diversity
shannon <- diversity(PCC.2018a, "shannon")

# Simpson's Evenness
simp.even <- apply(PCC.2018a, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(PCC.2018a[,-c(1:1)]))

#combined richness, diversity, evenness
diversity.plant.even <- cbind(meta,richness,shannon,simp.even,J)

#Adding diversity fields to dataset
meta$plant.shannon <- diversity.plant.even$shannon
meta$plant.J <- diversity.plant.even$J
meta$plant.simp <- diversity.plant.even$simp.even
meta$plant.rich <- diversity.plant.even$richness
colnames(meta) #total columns is 36

```

# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Plant, echo=TRUE}
library(lmerTest)

plant.rich.lm <- aov(plant.rich ~ treatment, data = meta)
plot(plant.rich.lm)
anova(plant.rich.lm)
#Response: plant.rich
 #         Df Sum Sq Mean Sq F value  Pr(>F)  
#treatment  1   60.5  60.500    13.2 0.01092 *
#Residuals  6   27.5   4.583  

plant.shannon.lm <- aov(plant.shannon ~ treatment, data = meta)
plot(plant.shannon.lm)
anova(plant.shannon.lm)
#Response: plant.shannon
 #         Df  Sum Sq  Mean Sq F value  Pr(>F)  
#treatment  1 0.28409 0.284095  9.6546 0.02092 *
#Residuals  6 0.17655 0.029426  
```

#Plot Plant Richness
```{r Plot - Plant Species Richness, echo=TRUE}
# Graphing Richness
p <- ggplot(meta, aes(x=treatment, y=plant.rich, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
shannon.plant <-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Plant Species Richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized"))
    #+ ggtitle("Plant Diversity (2018 - wet)")
shannon.plant
ggsave("../figures/2018.plant.richness.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#Plot shannon diversity 
```{r Plot - Plant Shannon Diversity, echo=TRUE}
# Graphing Shannon Diversity
p <- ggplot(meta, aes(x=treatment, y=plant.shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
shannon.plant <-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized"))+ ggtitle("Plant Diversity (2018 - wet)")
shannon.plant
ggsave("../figures/2018.plant.shannon.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# Simple Hypothesis Testing - Plants
```{r perMANOVA - Plant, echo=TRUE}
#PERMANOVA
adonis = adonis(PCC.2018.even[,-c(1:3)]~treatment, method = "bray", data = PCC.2018.even, perm=1000)
adonis
```
# Plant Ordinations
```{r Ordination (PCoA) - Plant, include=FALSE}
# Principal Coordinates Analysis
dataREL.dist <- vegdist(PCC.2018.even[,-c(1:3)], method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #54.3
explainvar2b #23.2
```

```{r Ordination (PCoA) - Plant, include=FALSE}
# Principal Coordinates Analysis - plants
pcoa.groups <- paste(PCC.2018.even$treatment)
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt <- c("M","MF")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt)

#Plot
df2a <- as.data.frame(pcoa.cent.dataframe.trts)
plot2a <- ggplot(df2a, aes(x=V1, y=V2, colour=trt)) + theme_bw() 
plot2a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  scale_colour_manual(labels = c("mowed/unfertilized","mowed/fertilized"), 
                      values = c("gray70", "darkgreen")) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (54.3%)") + ylab("PCoA 2 (23.2%)") +
        labs(colour = "Treatment") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4))) 
    #+ ggtitle("Plant Community Composition (2018 - wet)")
ggsave("../figures/2018.plant.even.ordination.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## Bacterial Data - mowed - even only
```{r read in 16S rRNA, include=FALSE}
#Read in OTU file
otu <- read.otu(sharedfile)
#Remove OTUs with less than 2 occurences across all sites

b <- ncol(otu)
reads <- sum(colSums(otu))
#reads before filtering 

otu.2 <- otu[, which(colSums(otu) > 10)]

#Remove all RP samples
#In the original WRC2015 sampling I collected bulk soils, and rhizospehre soils. Soil loosley adhering to roots was shaken into a bag (RP) for DNA and soil NH4/NO3 extraction. Roots and adhering soils collected for DNA processing (RS). Removed RP from DNA analysis.
#WRC15_65-WRC15_80
rownames(otu.2)
otu.2 <-otu.2[-c(33:48),]
rownames(otu.2)

#FROM OTU - remove blocks 1, 3, 5, 7 and treatments C & F
#First join meta and otu. Filter and then remove extra columns
meta_all <- read.csv(metafile)

otu.2.filter <- cbind(meta_all, otu.2)

otu.2.filter <- otu.2.filter %>%
  filter(year==2018) %>%
  filter(source=="bulk") %>%
  filter(treatment=="M" | treatment =="MF") %>%
  filter(block==2 | block==4 | block==6 | block==8)

rownames(otu.2.filter)<- otu.2.filter$sample

otu.2 <- otu.2.filter[,-c(1:26)]
rownames(otu.2.filter)
colnames(otu.2.filter)

#What sample has lowest read count? WRC18_201 16668, next lowest is WRC18_200 43192 highest WRC15_167 68819
otu2<- colSums(t(otu.2))
otu2
otu2[which.min(colSums(t(otu.2)))]
otu2[which.max(colSums(t(otu.2)))]

# Import Taxonomy File
otu.tax <- read.tax(taxonomy = "../data/WRC18_opti_mcc.0.03.cons.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
```

```{r sequence read counts by sample}
#graph of read counts, need to work on y axis
c<-as.data.frame(otu.2)
c$colsum <- colSums(t(c))
c$group <- rownames(c)
p<-ggplot(c, aes(x=group,y=colsum))+geom_bar(stat="identity")+coord_flip()
p

#Rarefy
#Set min sample number
min.N <- min(rowSums(otu.2))
#rrarefy returns a randomly rarefied community data frame or vector of selected size
#This is the otu set that should be used for beta diversity metrics
otu.rare <- rrarefy(otu.2, min.N)

```

```{r rarefaction curve}
#rarefy returns a species richness estimate in a random subsample of selected size
otu.rare.rich <- rarefy(x = otu.2, sample=min.N, se=T)
#Plot rarefaction curve
###suggests that we did not capture all of the diversity b/c the accumulation curve does not level off or flatten - rather, it continues to increase
rarecurve(x = otu.2, step = 20, col="blue", cex=0.6, las=1)
abline(0,1,col='red')
text(1000,1000,"1:1",pos=2, col='red')
```

```{r Relative abundance and PERMANOVA}
#transforms read counts to relative abundance
#Using rarefied OTU dataset
otu.2.rel <- otu.rare
for(i in 1:dim(otu.rare)[1]){
  otu.2.rel[i,] <- otu.rare[i,]/sum(otu.rare[i,])
}

#Combine experimental deisng meta data with OTU relative abundance data
metaotu <- cbind(meta, otu.2.rel)
```

# Diversity Metrics - Hypothesis Testing
```{r Bac Presence_Absence and Species richness, include =FALSE }
#Transform data to reflect only presence or absence ie 0 or 1
otu.2.PA<-(otu.rare>0*1)
#sanity check 
#otu.2.PA[,1:5]
#Add meta data
metaotu.2 <- cbind(meta, otu.2.PA)
#metaotu.2[1:15]
#Transform True/False to 0/1 
cols <- sapply(metaotu.2, is.logical)
metaotu.2[,cols]<-lapply(metaotu.2[,cols],as.numeric)
#head(metaotu.2)
```

```{r diversity metrics - calculate diversity and evenness}
# Calculate diversity metrics
bac.shannon <- vegan::diversity(otu.rare, "shannon")
bac.simp <- vegan::diversity(otu.rare, "simpson")
bac.invsimp <-vegan::diversity(otu.rare, "invsimpson")

# Pielous evenness (uses presence/absence data)
bac.J <- bac.shannon/log(specnumber(otu.2.PA[,-c(1:1)]))

#OTU richness
bac.richness <- rowSums((otu.2.PA >= 1))

#Adding diversity fields to dataset
meta$bac.shannon <- bac.shannon
meta$bac.J <- bac.J
meta$bac.simp <- bac.simp
meta$bac.invsimp <- bac.invsimp
meta$bac.richness <- bac.richness
colnames(meta) #total non-OTU columns is 42
```

# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Bacteria, echo=TRUE}
bac.shannon.lm <- aov(bac.shannon ~ treatment, data = meta)
plot(bac.shannon.lm)
anova(bac.shannon.lm)
#Response: bac.shannon
#          Df  Sum Sq  Mean Sq F value  Pr(>F)  
#treatment  1 0.25958 0.259579  7.9502 0.03037 *
#Residuals  6 0.19590 0.032651     
```

#Plot shannon diversity - bacteria
```{r Plot - Bacteria Shannon Diversity, echo=TRUE}
# Graphing Shannon Diversity
p <- ggplot(meta, aes(x=treatment, y=bac.shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
bac.shannon<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized")) + ggtitle("Bacterial Diversity (2018 - wet)")
bac.shannon
ggsave("../figures/2018.bacteria.even.shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Bacteria. Richness, echo=TRUE}
bac.richness.lm <- aov(bac.richness ~ treatment, data = meta)
plot(bac.richness.lm)
anova(bac.richness.lm)
#Response: bac.richness
#          Df Sum Sq Mean Sq F value  Pr(>F)  
#treatment  1 361250  361250  3.9639 0.09359 .
#Residuals  6 546808   91135     
```

#Plot OTU richness - bacteria
```{r Plot - Bacterial OTU richness, echo=TRUE}
# Graphing Richness Diversity
p <- ggplot(meta, aes(x=treatment, y=bac.richness, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
bac.shannon<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Bacterial OTU Richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized")) 
    #+ ggtitle("Bacterial OTU Richness (2018 - wet)")
bac.shannon
ggsave("../figures/2018.bacteria.even.richness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Bacterial Ordinations
```{r Ordination (PCoA) - Bacteria, include=FALSE}
#PERMANOVA
metaotu.ad <- adonis(metaotu[,-c(1:42)] ~ treatment, method = "bray", data = metaotu, perm=1000, set.seed=42)
metaotu.ad 

# Principal Coordinates Analysis
dataREL.dist <- vegdist(metaotu[,-c(1:42)], method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #37.4
explainvar2b #23.2
```

```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Principal Coordinates Analysis
pcoa.groups <- paste(metaotu$treatment)
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt <- c("M","MF")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt)

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=trt)) + theme_bw() 
bact.ord <-plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
  theme(panel.background = element_blank()) + 
  geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + 
  geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
  geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
  scale_colour_manual(labels = c("mowed/unfertilized","mowed/fertilized"), 
                      values = c("gray70", "darkgreen")) +
  theme(axis.title = element_text(size=14), 
        axis.text = element_text(size=14),
        axis.text.x = element_text(size=14), 
        panel.border = element_rect(colour = "black", size = 1.25)) + 
  theme(axis.ticks.length = unit(0.3, "cm")) + 
        xlab("PCoA 1 (37.4%)") + ylab("PCoA 2 (23.2%)") +
        labs(colour = "Treatment") +
    guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
           shape = guide_legend(override.aes = list(size = 4))) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized")) 
    #+ ggtitle("Bacterial Community Composition (2018 - wet)")
bact.ord
ggsave("../figures/2018.bact.ordination.bulk.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# CORRELATIONS Plant - Bacteria - wet (block=even)
```{r mantel test, include=FALSE}
colnames(metaotu)  #total non-OTU columns is 42

dist.plant <- vegdist(PCC.2018.even[,-c(1:3)], method="bray")
dist.bac <- vegdist(metaotu[,-c(1:42)], method="bray")

#MANTEL - all even blocks
mantel.rtest(dist.plant, dist.bac, nrepet = 999) #all treatments and ditch included - Mantel r = 0.631, p = 0.006

#separate by treatment

##PLANT
PCC.2018.M <- subset(PCC.2018.even, treatment == "M")
PCC.2018.MF <- subset(PCC.2018.even, treatment == "MF")
dist.plant.M <- vegdist(PCC.2018.M[,-c(1:3)], method="bray")
dist.plant.MF <- vegdist(PCC.2018.MF[,-c(1:3)], method="bray")

##BACTERIA
bac.2018.M <- subset(metaotu, treatment == "M")
bac.2018.MF <- subset(metaotu, treatment == "MF")
dist.bac.M <- vegdist(bac.2018.M[,-c(1:42)], method="bray")
dist.bac.MF <- vegdist(bac.2018.MF[,-c(1:42)], method="bray")

# MANTEL TESTS - even blocks only 
mantel.rtest(dist.plant, dist.bac, nrepet = 999) # mowed only: Mantel r = 0.631, p = 0.004
mantel.rtest(dist.plant.M, dist.bac.M, nrepet = 999) # mowed only: Mantel r = 0.564, p = 0.159
mantel.rtest(dist.plant.MF, dist.bac.MF, nrepet = 999) # mowed/fert: Mantel r = 0.369, p = 0.287

# correlation alpha diversity 
colnames(meta)
meta.M <- subset(meta, treatment == "M")
meta.MF <- subset(meta, treatment == "MF")

cor.test(meta$plant.shannon, meta$bac.shannon) # ALL: Pearsonn rho = -0.940, p = 0.0008
cor.test(meta.M$plant.shannon, meta.M$bac.shannon) # ALL: Pearsonn rho = -0.503, p = 0.497
cor.test(meta.MF$plant.shannon, meta.MF$bac.shannon) # ALL: Pearsonn rho = -0.940, p = 0.060

p <- ggplot(meta, aes(x=plant.shannon, y=bac.shannon, color=treatment)) +
  scale_color_manual(name="Treatment", values=c("gray70","darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) +  
  geom_point(size=4)
p1=p+geom_smooth(method="lm")
shannon.cor <- p1 + theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(axis.title=element_text(vjust=1,size=16), 
        axis.text=element_text(size=16), 
        axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
        panel.border = element_rect(colour = "black",size=1)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  labs(x = "Plant Shannon Diversity Index (H')", y = "Bacterial Shannon Diversity Index (H')") +
  theme(strip.text.x = element_text(size=16, face="bold"), strip.text.y = element_text(size=16, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1)) #+ ggtitle("Bacterial-Plant Relationship (2018 - wet)")
shannon.cor
ggsave("../figures/2018.plant.bact.shannon.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
# CORRELATIONS Plant - Fungal - wet (block=even)
```{r mantel test, include=FALSE}

##FUNGAL
#fun_metaotu <- cbind(meta, fungal_otu.rel)
#fun.2018.otu <- subset(fun_metaotu[,-c(1:26)])

##separate M vs MF for Mantel tests later
#fun.2018.metaotu.M <- subset(fun_metaotu, treatment == "M")
#fun.2018.metaotu.MF <- subset(fun_metaotu, treatment == "MF")
#fun.2018.otu.M <- subset(fun.2018.metaotu.M[,-c(1:26)])
#fun.2018.otu.MF <- subset(fun.2018.metaotu.MF[,-c(1:26)])

dist.fungal <- vegdist(fun.2018.otu, method="bray")
dist.fungal.M <- vegdist(fun.2018.otu.M, method="bray")
dist.fungal.MF <- vegdist(fun.2018.otu.MF, method="bray")

##PLANT - completed in previous chunk
#dist.plant <- vegdist(PCC.2018[,-c(1:3)], method="bray")
#dist.plant.M <- vegdist(PCC.2018.M[,-c(1:3)], method="bray")
#dist.plant.MF <- vegdist(PCC.2018.MF[,-c(1:3)], method="bray")

# MANTEL TESTS
mantel.rtest(dist.plant, dist.fungal, nrepet = 999) #ALL - Mantel r = 0.664, p = 0.002
mantel.rtest(dist.plant.M, dist.fungal.M, nrepet = 999) #mowed only: Mantel r = 0.405, p = 0.242
mantel.rtest(dist.plant.MF, dist.fungal.MF, nrepet = 999) #mowed/fert: Mantel r = 0.569, p = 0.071 

# correlation shannon diversity
cor.test(meta$fun.shannon, meta$plant.shannon) # ALL: Pearson rho = 0.185, p = 0.661
cor.test(meta.M$fun.shannon, meta.M$plant.shannon) # mowed only: Pearson rho = -0.069, p = 0.932
cor.test(meta.MF$fun.shannon, meta.MF$plant.shannon) # mowed/fert: Pearson rho = -0.268, p = 0.732
 
# correlation shannon diversity
cor.test(meta.M$fun.shannon, meta.M$bac.shannon) # mowed only: Pearson rho = 0.921, p = 0.079
cor.test(meta.MF$fun.shannon, meta.MF$bac.shannon) # mowed/fert: Pearson rho = 0.433, p = 0.567

# change the x and y to reflect plant and fungi
p <- ggplot(meta, aes(x=plant.shannon, y=fun.shannon, color=treatment)) +
  scale_color_manual(name="Treatment", values=c("gray70","darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) +  
  geom_point(size=4)
p1=p+geom_smooth(method="lm")
shannon.cor <- p1 + theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(axis.title=element_text(vjust=1,size=16), 
        axis.text=element_text(size=16), 
        axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
        panel.border = element_rect(colour = "black",size=1)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  labs(x = "Plant Shannon Diversity Index (H')", y = "Fungal Shannon Diversity Index (H')") +
  theme(strip.text.x = element_text(size=16, face="bold"), strip.text.y = element_text(size=16, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1)) #+ ggtitle("Plant-Fungal Relationship (2018 - wet)")
shannon.cor
ggsave("../figures/2018.plant.fun.shannon.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

# change the x and y to reflect plant and fungi
p <- ggplot(meta, aes(x=fun.shannon, y=bac.shannon, color=treatment)) +
  scale_color_manual(name="Treatment", values=c("gray70","darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) +  
  geom_point(size=4)
p1=p+geom_smooth(method="lm")
shannon.cor <- p1 + theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(axis.title=element_text(vjust=1,size=16), 
        axis.text=element_text(size=16), 
        axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
        panel.border = element_rect(colour = "black",size=1)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  labs(x = "Fungal Shannon Diversity Index (H')", y = "Bacterial Shannon Diversity Index (H')") +
  theme(strip.text.x = element_text(size=16, face="bold"), strip.text.y = element_text(size=16, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1)) #+ ggtitle("Bacterial-Fungal Relationship (2018 - wet)")
shannon.cor
ggsave("../figures/2018.bact.fun.shannon.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

```


## carbon metabolic profile
```{r}
file_path <- ("../data/2018WRC_CNPS_shannon.csv")
CNPS_shannon <- read.csv(file_path)
CNPS_shannon_bulk <- subset(CNPS_shannon, source=='bulk')
meta$eco_shannon <- CNPS_shannon_bulk$eco_shannon

meta.M <- subset(meta, treatment == "M")
meta.MF <- subset(meta, treatment == "MF")

cor.test(meta$fun.shannon, meta$eco_shannon) # ALL: Pearson rho = 0.075, p = 0.860
cor.test(meta$plant.shannon, meta$eco_shannon) # ALL: Pearson rho = -0.023, p = 0.957
cor.test(meta$bac.shannon, meta$eco_shannon) # ALL: Pearson rho = -0.015, p = 0.972

cor.test(meta.M$bac.shannon, meta.M$eco_shannon) # mowed only: Pearson rho = 0.400, p = 0.600
cor.test(meta.MF$bac.shannon, meta.MF$eco_shannon) # mowed/fert: Pearson rho = 0.621, p = 0.379


p <- ggplot(meta, aes(x=eco_shannon, y=bac.shannon, color=treatment)) +
  scale_color_manual(name="Treatment", values=c("gray70","darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) +  
  geom_point(size=4)
p1=p+geom_smooth(method="lm")
shannon.cor <- p1 + theme_bw() + 
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) + 
  theme(axis.title=element_text(vjust=1,size=16), 
        axis.text=element_text(size=16), 
        axis.text.x = element_text(vjust=0.65, hjust=0.5, size=16), 
        panel.border = element_rect(colour = "black",size=1)) +
  theme(axis.ticks.length=unit(0.3,"cm")) + 
  labs(x = "Carbon Metabolic Diversity Index (H')", y = "Bacterial Shannon Diversity Index (H')") +
  theme(strip.text.x = element_text(size=16, face="bold"), strip.text.y = element_text(size=16, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1)) #+ ggtitle("Bacterial-Carbon Process Relationship (2018 - wet)")
shannon.cor
ggsave("../figures/2018.bact.carbon.shannon.even.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#Plot carbon metabolism
```{r Plot - carbon metabolic diversity, echo=TRUE}
# Graphing carbon metabolic Shannon Diversity
carbon.diversity.lm <- aov(eco_shannon ~ treatment, data = meta)
plot(carbon.diversity.lm)
anova(carbon.diversity.lm) #NS

p <- ggplot(meta, aes(x=treatment, y=eco_shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("mowed/unfertilized", "mowed/fertilized")) 
p1=p+geom_smooth(method="lm")
carbon.shannon<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Treatment", y = "Carbon Metabolic Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("M", "MF"), labels=c("unfertilized", "fertilized")) + ggtitle("Carbon Metabolic Diversity (2018 - wet)")
carbon.shannon
ggsave("../figures/2018.carbon.metabolic.shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```