---
title: Distinct microbial communities alter litter decomposition rates in a fertilized coastal plain wetland
author: "Megan E. Koceja, Regina Bledsoe, Carol Goodwillie, Ariane L. Peralta"
date: "Last updated on `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
  fig_caption: yes
  pdf_document: null
header-includes:
- \usepackage{array}
- \usepackage{graphics}
- \usepackage[utf8]{inputenc}
---

Project Description: Fill out

# Initial Setup
```{r Initial Setup, include=FALSE}
rm(list=ls())
setwd("~/GitHub/WRC18_RhizoTeaDecomp/analyses")
se <- function(x, ...){sd(x, na.rm = TRUE)/sqrt(length(na.omit(x)))}
ci <- function(x, ...){1.96 * sd(x,na.rm = TRUE)}

# Code Dependencies
source("../bin/DiversityFunctions.R")
source("../bin/MothurTools.R")
require("vegan"); require("lme4"); require("reshape"); require("ggplot2")
require("ade4"); require("png"); require ("agricolae"); require("tidyr")
require("dplyr"); require("reshape2"); require("picante"); require("phyloseq")
```
#Import Files
## Environmental Data
```{r Import Files - Env, include=FALSE}
# Import Environmental Data
design <- read.csv("../data/2018_WRC_SoilMicroDecomp_missing3t.csv", row.names=1)
```
## Bacterial Data - including tea only
```{r Import Files - Bacteria, include=FALSE}
# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/WRC18_opti_mcc.shared")

# Correct Sample IDs and Subset File
missing <- setdiff(rownames(otu.in), rownames(design))
otu <- otu.in[-(which(rownames(otu.in) %in% missing)), ]
dim(otu)
otu <- otu[match(rownames(design), rownames(otu)),]
all.equal(rownames(otu), rownames(design))

# OTU table - remove otus w/ < 2 occurrences across all sites
otu_keep <- otu[, which(colSums(otu) >= 2)]
dim(otu_keep)

#otus <- otu[, which(colSums(otu) >= 2)]

aa <- (rowSums(otu_keep))
aa # WRC18_10_Tea-155 reads WRC18_13_Tea-29 reads - removed

# OTU table - removed low abundance samples
otu_low_remov <- otu_keep[which(rowSums(otu_keep) >= 1000), ]
dim(otu_low_remov)

# OTU table - odd sites in bacterial composition data and remove in design file
odd.sites <- c("WRC18_10_Tea","WRC18_13_Tea")

otu_final <- otu_low_remov[setdiff(rownames(otu_low_remov), odd.sites), ]
design_final <- design[setdiff(rownames(design), odd.sites), ]
all.equal(rownames(design_final), rownames(otu_final))

otus <- otu_final

###subset tea only

otus.design <- cbind(design_final,otus)
otus.design.teas <- subset(otus.design, source == "green_tea" | source == "rooibos_tea")
dim(otus.design.teas)                   

design.t <- otus.design.teas[,c(1:16)]
otus.t <- otus.design.teas[,-c(1:16)]
#full df - otus.design.teas

# Make Relative Abundance Matrices
dataREL.t <- otus.t
for(i in 1:dim(otus.t)[1]){
  dataREL.t[i,] <- otus.t[i,]/sum(otus.t[i,])
}
```
# Simple Hypothesis Testing - Microbial Community (tea only)
```{r perMANOVA - Bacteria - All Yrs, echo=TRUE}
#PERMANOVA
design.t$ditch <- as.factor(design.t$ditch)
new.data.t <-cbind(design.t,dataREL.t)
adonis = adonis(new.data.t[,-c(1:16)]~source*treatment*ditch, method = "bray", data = new.data.t, perm=1000, set.seed=42)
adonis
```
# str-fxn tea type only
```{r perMANOVA - Bacteria - All Yrs, echo=TRUE}
# bcc = str(new.data.t)
#Redundancy Analysis
mod1 <- rda(dataREL.t ~ design.t$Percent_Loss)
mod1
plot(mod1)

f <- dataREL.t ~ design.t$Percent_Loss
df.rda <- dbrda(f, data=dataREL.t, distance="bray")
anova(df.rda, by="terms", model="direct")
RsquareAdj(df.rda)

#new.data.t <-cbind(design.t,dataREL.t)

library(dbstats)
require(pls)
library(pls)
library(plsr)

df <- vegdist(dataREL.t, method="bray")

decomp.dbplsr <- dbplsr(new.data.t$Percent_Loss ~ as.matrix(df), data=new.data.t, ncomp=6, method="GCV")
summary(decomp.dbplsr)
```
# Bacterial Ordinations

## Principal Coordinates Ordination
```{r Ordination (PCoA) - Bacteria - tea only, include=FALSE}
# Principal Coordinates Analysis
dataREL.dist <- vegdist(dataREL.t, method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #22.2
explainvar2b #10.1
```

```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Principal Coordinates Analysis - teas only #NEED to incorporate ditch
new.data.t <-cbind(design.t,dataREL.t)
droplevels(new.data.t$source)
pcoa.groups <- paste(new.data.t$treatment, new.data.t$source, new.data.t$ditch, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt.t <- c("M","M","M","M","MF","MF","MF","MF")
source.ditch.t <- c("green_tea (wet_0ditch)","green_tea (dry_1ditch)","rooibos_tea (wet_0ditch)","rooibos_tea (dry_1ditch)","green_tea (wet_0ditch)","green_tea (dry_1ditch)","rooibos_tea (wet_0ditch)","rooibos_tea (dry_1ditch)")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt.t)
pcoa.cent.dataframe.trts$source <- as.factor(source.ditch.t)

#pcoa.col <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 2)) # group
#pcoa.shape <- as.factor(sapply(strsplit(pcoa.cent.treats, "_"), `[`, 1)) # group

#Plot
df1a <- as.data.frame(pcoa.cent.dataframe.trts)
plot1a <- ggplot(df1a, aes(x=V1, y=V2, colour=trt, shape = source,
                           group = interaction(trt,source))) + theme_bw() 
plot1a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
theme(panel.background = element_blank()) + 
geom_point(aes(colour=trt), size=5, stroke = 1.25, show.legend = TRUE) +
geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
scale_colour_manual(labels = c("unfertilized","fertilized"), 
                    values = c("gray70", "darkgreen")) + 
scale_shape_manual(labels = c("green_tea (dry_1ditch)", 
                              "green_tea (wet_0ditch)", 
                              "rooibos_tea (dry_1ditch)",
                              "rooibos_tea (wet_0ditch)"), 
                   values = c(21,16,24,17)) +
theme(axis.title = element_text(size=14), axis.text = element_text(size=14),
      axis.text.x = element_text(size=14), 
      panel.border = element_rect(colour = "black", size = 1.25)) + 
theme(axis.ticks.length = unit(0.3, "cm")) + 
xlab("PCoA 1 (22.2%)") + ylab("PCoA 2 (10.1%)") +
  labs(colour = "Treatment", shape = "Source + Ditch") +
  guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
         shape = guide_legend(override.aes = list(size = 4)))
ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/Ordination_tea.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

#Loading data file
```{r Loading Files, include=FALSE}
# load data file for TBI
TBI <- read.csv("../data/2018_WRC_TBI_Rstats.csv", row.names=1) #removed sampleID#17
dim(TBI)
```

# relationship between S and k
```{r Loading Files, include=FALSE}
TBI$Ditch_Trt <- as.factor(TBI$Ditch_Trt)
TBI$ditch <- as.factor(TBI$Ditch)
TBI$trt <- as.factor(TBI$Treatment)

#different shapes
p <- ggplot(TBI, aes(x=S, y=k, color=Ditch_Trt, shape=Ditch_Trt))+ 
        scale_color_manual(name="Ditch + Treatment", 
                           values=c("gray70","darkgreen","gray70","darkgreen"), 
                            labels = c("wet_0ditch + unfertilized",
                                      "wet_0ditch + fertilized",
                                      "dry_1ditch + unfertilized",
                                      "dry_1ditch + fertilized")) +
  geom_point(size=5) + 
        scale_shape_manual(name="Ditch + Treatment", values = c(18,16,23,21), 
                           labels = c("wet_0ditch + unfertilized",
                                      "wet_0ditch + fertilized",
                                      "dry_1ditch + unfertilized",
                                      "dry_1ditch + fertilized"))
p1=p+geom_point()
TBI <- p1 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +theme(axis.title=element_text(vjust=1,size=18), axis.text=element_text(size=18), axis.text.x = element_text(vjust=0.65, hjust=0.5, size=18), panel.border = element_rect(colour = "black",size=1)) + theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Stabilization factor (S)", y = "Decomposition rate (per day) (k)") + theme(strip.text.x = element_text(size=18, face="bold"), strip.text.y = element_text(size=18, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1))
TBI

ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/TBI.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#same shape
p <- ggplot(TBI, aes(x=S, y=k, color=trt, shape=ditch))+ scale_color_manual(name="Treatment", values=c("gray70","darkgreen"), labels = c("unfertilized", "fertilized")) + geom_point(size=5) + scale_shape_manual(name="Ditch", values = c(16,21), labels = c("wet_0ditch","dry_1ditch"))
p1=p+geom_point()
TBI <- p1 + theme_bw() + theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line = element_line(colour = "black")) +theme(axis.title=element_text(vjust=1,size=18), axis.text=element_text(size=18), axis.text.x = element_text(vjust=0.65, hjust=0.5, size=18), panel.border = element_rect(colour = "black",size=1)) + theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Stabilization factor (S)", y = "Decomposition rate (per day) (k)") + theme(strip.text.x = element_text(size=18, face="bold"), strip.text.y = element_text(size=18, face="bold"), strip.background = element_rect(colour="black", fill="white", size=1))
TBI
```

# Tea Decomp Biomass
```{r Loading Files, include=FALSE}
library("lmerTest")
library("pbkrtest")

#subset tea only rows
tea <- subset(design, source != "bulk_soil")
tea$ditch <- as.factor(tea$ditch)
str(tea)

mass_loss.lm <- lmer(Percent_Loss ~ source*treatment*ditch + (1|block), tea)
plot(mass_loss.lm)
mass_loss.lm
anova(mass_loss.lm, type="II", ddf="Kenward-Roger")

#summary table for mass loss
summary <- tea %>% group_by(source, treatment, ditch) %>% summarise(mean.biomass=mean(Percent_Loss), se.biomass=se(Percent_Loss))
print(summary)
write.csv(summary,"../data/massloss.summary.csv")

#aov.fit2 <- aov(Percent_Loss~treatment*source*ditch, tea)
#summary(aov.fit2)

#posthoc2 <- TukeyHSD(x=aov.fit2, conf.level=0.95)
#posthoc2

# Graphing Percent Loss 
ditch.labs <- c("wet plots farther from ditch","dry plots adjacent to ditch") #for facet labels
names(ditch.labs) <- c("0", "1")

p <- ggplot(tea, aes(x=source, y=Percent_Loss, color=as.factor(treatment)))+ scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("unfertilized", "fertilized")) + stat_summary(fun.data=mean_cl_boot,size=0.75) 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Source", y = "Percent Loss") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("green_tea", "rooibos_tea"), labels=c("green tea", "rooibos tea"))
ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/PercentLoss_pointCI.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)

#boxplots
p <- ggplot(tea, aes(x=source, y=Percent_Loss, color=as.factor(treatment))) + geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) +   
      scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("unfertilized", 
      "fertilized")) +
    labs(tag = "Source: P<0.001\nTreatment: P<0.001\nSource*Trt: P=0.045") 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Source", y = "Percent Loss") + 
    theme(strip.text.x = element_text(size=13, face="bold"), strip.text.y =   
          element_text(size=12, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("green_tea", "rooibos_tea"), labels=c("green tea", "rooibos tea"))+
     theme(plot.tag.position = c(1.01, .85)) + theme(plot.tag=element_text(hjust=1,size=10))

ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/PercentLoss_boxplots.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

## Bacterial Data - including bulk and tea
```{r Import Files - Bacteria, include=FALSE}
# Important Environmental data
design <- read.csv("../data/2018_WRC_SoilMicroDecomp_missing3t.csv", row.names=1)

# Import OTU data
# Import Raw Data
otu.in <- read.otu("../data/WRC18_opti_mcc.shared")

# Correct Sample IDs and Subset File
missing <- setdiff(rownames(otu.in), rownames(design))
otu <- otu.in[-(which(rownames(otu.in) %in% missing)), ]
dim(otu)
otu <- otu[match(rownames(design), rownames(otu)),]
all.equal(rownames(otu), rownames(design))

# OTU table - remove otus w/ < 2 occurrences across all sites
otu_removal <- otu[, which(colSums(otu) >= 2)]
dim(otu_removal)

aa <- (rowSums(otu_removal))
aa # WRC18_10_Tea-155 reads WRC18_13_Tea-29 reads - removed

# OTU table - removed low abundance samples
otu_low_remov <- otu_removal[which(rowSums(otu_removal) >= 1000), ]
dim(otu_low_remov)

# OTU table - odd sites in bacterial composition data and remove in design file
odd.sites <- c("WRC18_10_Tea","WRC18_13_Tea")

otu_final <- otu_low_remov[setdiff(rownames(otu_low_remov), odd.sites), ]
design_final <- design[setdiff(rownames(design), odd.sites), ]
all.equal(rownames(design_final), rownames(otu_final))

otus <- otu_final

# Make Presence Absence Matrix
dataPA <- (otus > 0) * 1

# Make Relative Abundance Matrices
dataREL <- otus
for(i in 1:dim(otus)[1]){
  dataREL[i,] <- otus[i,]/sum(otus[i,])
}

# Import Taxonomy File
otu.tax <- read.tax(taxonomy = "../data/WRC18_opti_mcc.0.03.cons.taxonomy",
                   format = "rdp", tax.levels = 6, col.tax = 3)
```
# Diversity Metrics - Hypothesis Testing
```{r Diversity Metrics - Bacteria, echo=TRUE}
# Rarefy Abundances (min abundance is 29485. We are sampling to 29000)
min(rowSums(otus))
otus.r <- rrarefy(otus, 29000)

# Fisher's Alpha
fisher <- fisher.alpha(otus.r)

# Species Richness
#richness <- rowSums((PWESdata.r >= 1))
richness <- rowSums((otus >= 1))

# Shannon Diversity
shannon <- diversity(otus.r, "shannon")

# Simpson's Evenness
simp.even <- apply(otus.r, 1, simp_even)

#Pielou’s evenness
J <- shannon/log(specnumber(otus.r[,-c(1:1)]))

#combined richness, diversity, evenness
diversity <- cbind(design_final,richness,shannon,simp.even,J)
write.csv(diversity,"../data/diversity.bact.raw.csv")
```

# Diversity Metrics - Hypothesis Testing
```{r Hypothesis Testing - Bacteria, echo=TRUE}
#summary table for bacterial diversity
print(summary)
write.csv(summary,"../data/diversity.bact.summary.csv")

#summary table for soils
summary <- diversity %>% group_by(source, treatment, ditch) %>% summarise(mean.soilC=mean(c_percent), se.soilC=se(c_percent), mean.soilN=mean(n_percent), se.soilN=se(n_percent),mean.soilCN=mean(C_N), se.soilCN=se(C_N),mean.ammonium=mean(n_nh4_ug_gsoil), se.ammonium=se(n_nh4_ug_gsoil),mean.nitrate=mean(n_no3_ug_gsoil), se.nitrate=se(n_no3_ug_gsoil),mean.phosphate=mean(p_po4_w_ug_gsoil), se.phosphate=se(p_po4_w_ug_gsoil))
print(summary)
write.csv(summary,"../data/diversity.soil.summary.csv")

library(lmerTest)
diversity$ditch <- as.factor(diversity$ditch)
str(diversity)

richness.lm <- lmer(richness ~ source*treatment*ditch + (1|block), data = diversity)
plot(richness.lm)
richness.lm
anova(richness.lm, type="II", ddf="Kenward-Roger")

evenness.lm <- lmer(simp.even ~ source*treatment*ditch + (1|block), data = diversity)
plot(evenness.lm)
evenness.lm
anova(evenness.lm, type="II", ddf="Kenward-Roger")

shannon.lm <- lmer(shannon ~ source*treatment*ditch + (1|block), data = diversity)
plot(shannon.lm)
shannon.lm
anova(shannon.lm, type="II", ddf="Kenward-Roger")

#regression alpha diversity
# Multiple Linear Regression Example 
diversity.teas <- subset(diversity, source == "green_tea" | source == "rooibos_tea")

fit <- lm(Percent_Loss ~ richness + shannon + simp.even, diversity.teas)
summary(fit) #
plot(fit)

```

#Plot shannon diversity 
```{r Plot - Shannon Diversity, echo=TRUE}
ditch.labs <- c("wet plots farther from ditch", "dry plots adjacent to ditch") #for facet labels
names(ditch.labs) <- c("0", "1")

# Graphing Shannon Diversity
p <- ggplot(diversity, aes(x=source, y=shannon, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("unfertilized", "fertilized")) + labs(tag = "Source: P<0.0001\nTreatment: P<0.0001\nDitch: NS") 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
shannon<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = "Shannon Diversity Index (H')") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("bulk_soil", "green_tea", "rooibos_tea"), labels=c("soil", "green", 
          "rooibos")) + 
    scale_y_continuous(limits =c(3.8,7.00),labels = scales::number_format(accuracy = 0.01)) +
     theme(plot.tag.position = c(.99, .88)) + theme(plot.tag=element_text(hjust=1,size=10))

shannon
#ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/shannon.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#Plot richness 
```{r Plot - OTU richness, echo=TRUE}
# Graphing Richness
p <- ggplot(diversity, aes(x=source, y=richness, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("unfertilized", "fertilized")) + labs(tag = "Source: P<0.0001\nTreatment: P=0.009\nDitch: NS") 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
richness<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "", y = "OTU Richness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("bulk_soil", "green_tea", "rooibos_tea"), labels=c("soil", "green", 
          "rooibos"))+
     theme(plot.tag.position = c(0.99, .77)) + theme(plot.tag=element_text(hjust=1,size=10))
richness
#ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/richness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

#Plot Simpson's Evenness
```{r Plot Simpsons Evenness, echo=TRUE}
# Graphing Simpson's Evenness
p <- ggplot(diversity, aes(x=source, y=simp.even, color=as.factor(treatment)))+ geom_boxplot() +
      geom_point(aes(color=factor(treatment)), size=2, position = position_jitterdodge()) + scale_color_manual(name="Treatment", values=c("gray70", "darkgreen"), labels = c("unfertilized", "fertilized")) + labs(tag = "Source: P=0.035\nTreatment: P=0.005\nDitch: NS") 
p1=p+geom_smooth(method="lm")+facet_wrap(~ditch)+facet_grid(. ~ ditch,labeller = labeller(ditch=ditch.labs))
even<-p1 + theme_bw() + 
    theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(), axis.line 
          =element_line(colour = "black")) + 
    theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(vjust=0.65, hjust=0.5, 
          size=14), panel.border = element_rect(colour = "black",size=1)) + 
    theme(axis.ticks.length=unit(0.3,"cm")) + labs(x = "Source", y = "Simpson's Evenness") + 
    theme(strip.text.x = element_text(size=14, face="bold"), strip.text.y =   
          element_text(size=14, face="bold"), strip.background = element_rect(colour="black",
          fill="white", size=1)) +
    scale_x_discrete(breaks=c("bulk_soil", "green_tea", "rooibos_tea"), labels=c("soil", "green", 
          "rooibos")) + 
    scale_y_continuous(limits =c(0.00,0.12),labels = scales::number_format(accuracy = 0.01)) +
     theme(plot.tag.position = c(.99, .88)) + theme(plot.tag=element_text(hjust=1,size=10))
even 
#ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/evenness.png", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```
#combine alpha diversity graphs
```{r Loading Files, include=FALSE}
library("cowplot") #combining plots
library("ggpubr")
# remove legends and axes
n1 <-richness + theme(legend.position="top", legend.box = "horizontal")
n2 <- shannon + theme(legend.position="none")
n3 <- even + theme(legend.position="none")

# align all plots vertically
plots <- align_plots(n1,n2,n3, align = 'v', axis = 'l')

# combine moisture N process graphs
n4 <- plot_grid(plots[[1]],plots[[2]],plots[[3]], labels = c("A","B","C"), label_size = 16, ncol=1, align='v')
n4 

ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/alpha_diversity.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1,  width=8, height=15, dpi=300, limitsize=TRUE)
```
# Simple Hypothesis Testing - Microbes
```{r perMANOVA - Bacteria - All sources, echo=TRUE}
#PERMANOVA
new.data <-cbind(design_final,dataREL)
adonis = adonis(new.data[,-c(1:16)]~source*treatment*ditch, method = "bray", data = new.data, perm=1000, set.seed=42)
adonis
```
# Bacterial Ordinations
```{r Ordination (PCoA) - Bacteria - bulk and teas, include=FALSE}
# Principal Coordinates Analysis
dataREL.dist <- vegdist(dataREL, method="bray")

pcoa <- cmdscale(dataREL.dist, k=3, eig=TRUE, add=FALSE)
  # Classical (Metric) Multidimensional Scaling; returns PCoA coordinates
  # eig=TRUE returns eigenvalues; k = # of dimensions to calculate

explainvar1b <- round(pcoa$eig[1] / sum(pcoa$eig), 3) * 100
explainvar2b <- round(pcoa$eig[2] / sum(pcoa$eig), 3) * 100
sum.eigb <- sum(explainvar1b, explainvar2b)

explainvar1b #43.3
explainvar2b #10.6
```

```{r Ordination (PCoA) - Bacteria, include=FALSE}
# Principal Coordinates Analysis - bulk and teas
new.data <-cbind(design_final,dataREL)
pcoa.groups <- paste(new.data$treatment, new.data$source, new.data$ditch, sep = "_")
pcoa.points <- data.frame(pcoa$points, group = pcoa.groups)

# Calculate Centroids (mean and SE)
pcoa.L.centroids <- melt(pcoa.points, id="group", measure.vars = c("X1", "X2"))
pcoa.centroids <- acast(pcoa.L.centroids, variable ~ group, mean)
pcoa.centroids.se <- acast(pcoa.L.centroids, variable ~ group, se)
pcoa.centroids.sd <- acast(pcoa.L.centroids, variable ~ group, sd)

# Combine
pcoa.cent.dataframe <- cbind(t(pcoa.centroids), t(pcoa.centroids.se))
colnames(pcoa.cent.dataframe) <- c("V1", "V2", "V1e", "V2e")
pcoa.cent.treats <- rownames(pcoa.cent.dataframe)
trt.t.s <- c("M","M","M","M","M","M","MF","MF","MF","MF","MF","MF")
source.ditch.t.s <- c("bulk_soil (wet_0ditch)","bulk_soil (dry_1ditch)","green_tea (wet_0ditch)","green_tea (dry_1ditch)","rooibos_tea (wet_0ditch)","rooibos_tea (dry_1ditch)","bulk_soil (wet_0ditch)","bulk_soil (dry_1ditch)","green_tea (wet_0ditch)","green_tea (dry_1ditch)","rooibos_tea (wet_0ditch)","rooibos_tea (dry_1ditch)")
pcoa.cent.dataframe.trts <- as.data.frame(pcoa.cent.dataframe)
pcoa.cent.dataframe.trts$trt <- as.factor(trt.t.s)
pcoa.cent.dataframe.trts$source <- as.factor(source.ditch.t.s)

#Plot
df2a <- as.data.frame(pcoa.cent.dataframe.trts)
plot2a <- ggplot(df2a, aes(x=V1, y=V2, colour=trt, shape = source,
                           group = interaction(trt, source))) + theme_bw() 
plot2a + theme(panel.grid.major = element_blank(), 
               panel.grid.minor = element_blank(), 
               axis.line = element_line(colour = "black")) + 
theme(panel.background = element_blank()) + 
geom_point(aes(colour=trt.t.s), size=5, stroke = 1.25, show.legend = TRUE) +
geom_errorbarh(aes(xmax=V1+V1e, xmin=V1-V1e, height=0.01), colour="black") + geom_errorbar(aes(ymax=V2+V2e, ymin=V2-V2e, width=0.01), colour="black") +
scale_colour_manual(labels = c("unfertilized","fertilized"), 
                    values = c("gray70", "darkgreen")) + 
scale_shape_manual(labels = c("bulk_soil (dry_1ditch)", 
                              "bulk_soil (wet_0ditch)",
                              "green_tea (dry_1ditch)", 
                              "green_tea (wet_0ditch)", 
                              "rooibos_tea (dry_1ditch)",
                              "rooibos_tea (wet_0ditch)"), 
                   values = c(22,15,21,16,24,17)) +
theme(axis.title = element_text(size=14), axis.text = element_text(size=14),
      axis.text.x = element_text(size=14), 
      panel.border = element_rect(colour = "black", size = 1.25)) + 
theme(axis.ticks.length = unit(0.3, "cm")) + 
xlab("PCoA 1 (43.3%)") + ylab("PCoA 2 (10.6%)") +
  labs(colour = "Treatment", shape = "Source + Ditch") +
  guides(colour = guide_legend(override.aes = list(pch=16, size = 4)),
         shape = guide_legend(override.aes = list(size = 4)))
ggsave("../figures/figs_2018_WRC_SoilTeaDecomp/Ordination_bulk_tea.jpg", plot=last_plot(), device=NULL, path=NULL, scale=1, width=7, height=5, dpi=300, limitsize=TRUE)
```

# Bacterial community indicator species
```{r Bacteria Indicator Species, include=FALSE}
#new.data <-cbind(design_final,dataREL) - includes bulk and tea sources
library("labdsv")
group = interaction(new.data$source, new.data$treatment, new.data$ditch)
design.type <- group #treatment x fertilization x ditch
dataREL.t <- new.data[,-c(1:16)]

dataREL <- dataREL.t[, colSums(dataREL.t) > 0.05]
bac.ind <- indval(dataREL, design.type)
levels(design.type)
summary(bac.ind)

inds <- which(bac.ind$pval <= 0.05)
bac.indicators <- as.data.frame(matrix(NA, nrow = length(inds), ncol = 4))
colnames(bac.indicators) <- c("OTU", "Cluster", "IndVal", "Prob")

bac.indicators$OTU <- names(inds)
bac.indicators$Cluster <- bac.ind$maxcls[inds]
bac.indicators$IndVal <- bac.ind$indcls[inds]
bac.indicators$Prob <- bac.ind$pval[inds]

ind.tax <- otu.tax[which(as.character(otu.tax$OTU) %in% bac.indicators$OTU), ]
ind.tax <- ind.tax[match(ind.tax$OTU, bac.indicators$OTU), ]

indicator.bac <- cbind(bac.indicators, ind.tax[, -c(1)])

indicator.bac <- indicator.bac[order(as.numeric(indicator.bac$Cluster)), ]

table(indicator.bac$Cluster)
table(indicator.bac$Phylum)
table(indicator.bac$Cluster)
levels(design.type)

# Export Bacteria Indicator Table
write.table(indicator.bac, "../data/BacterialIndicators_source.txt",
            sep="\t", row.names = F, quote = F)
```

```{r load taxonomy data into phyloseq object}
# Import otu and tax into phyloseq
otu <- otu_table(t(otus), taxa_are_rows = T)
taxfile="../data/WRC18_opti_mcc.0.03.cons.taxonomy"
tax <- import_mothur(mothur_constaxonomy_file =  taxfile)
#rownames(meta) <- meta$group
sample <- sample_data(design_final)
#import into phyloseq
phy<-phyloseq(otu,tax)
row.names(sample)=sample_names(phy)
#sample_names(sample)
#sample_names(phy)
#Assign rownames to be Sample ID's
# Merge mothurdata object with sample metadata
meta_merge <- merge_phyloseq(phy,sample)
#assign column names
colnames(tax_table(meta_merge))
colnames(tax_table(meta_merge)) <- c("Kingdom", "Phylum", "Class", "Order", "Family", "Genus")
colnames(tax_table(meta_merge))
rank_names(meta_merge)
```

```{r phyloseq Phylum stacked graph of taxa composition}

# melt to long format (for ggploting) 
# prune out phyla below 2% in each sample
wrc_phylum <- meta_merge %>%
  tax_glom(taxrank = "Phylum") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, source!="bulk_soil") %>%                        # Filter out low abundance taxa
  arrange(Phylum)                                      # Sort data frame alphabetically by phylum

# Set colors for plotting
phylum_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_phylum$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_phylum, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_phylum, aes(x = Phylum, y = Abundance, fill = i)) + 
  facet_wrap(~treatment) +
  geom_bar(stat = "summary", fun.y = "mean") +
  scale_fill_manual(values = phylum_colors) +
  # Remove x axis title
  theme_bw()+
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) +
  xlab ("Phylum") +
  ylab("Relative Abundance > 1.0% \n") +
  ggtitle("Phylum Composition of Bacterial Communities")

wrc_plot + coord_flip()

p <- aov(Abundance ~ source*treatment*ditch + (1|block), wrc_phylum)
plot(p)
anova(p)

wrc_phylum_means <- wrc_phylum %>%
  group_by(source, Phylum) %>%
  summarise_at(c("Abundance"), mean, na.rm=TRUE)

write.csv(wrc_phylum, file="../figures/figs_2018_WRC_SoilTeaDecomp/phylum_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/phylum_stack.pdf", width=9, height=9, dpi=300)
```

```{r phyloseq Class stacked graphs of taxa composition}
# melt to long format (for ggploting) 
# prune out phyla below 2% in each sample
wrc_class <- meta_merge %>%
  tax_glom(taxrank = "Class") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, source!="bulk_soil") %>%    # Filter out low abundance taxa
  arrange(Class)                                     # Sort data frame alphabetically by phylum

# Set colors for plotting
class_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_class$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_class, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_class, aes(x = Class, y = Abundance, fill = i)) + 
  facet_grid(.~treatment)+#, labeller=labeller(ditch=labels)) +
  geom_bar(stat = "summary", fun.y = "mean") + 
  scale_fill_manual(values = class_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  ylab("Class Relative Abundance > 1.0%") +
  xlab("Treatment") +
  labs(fill="Treatment") +
  ggtitle("Class Composition of Bacterial Communities") 

wrc_plot + coord_flip() 

p <- aov(Abundance ~ source*treatment*ditch + (1|block), wrc_class)
plot(p)
anova(p)

wrc_class_means <- wrc_class %>%
  group_by(source, Class) %>%
  summarise_at(c("Abundance"), mean, na.rm=TRUE)

write.csv(wrc_class_means, file="../figures/figs_2018_WRC_SoilTeaDecomp/class_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/class_stack.pdf", width=9, height=9, dpi=300)

```

```{r phyloseq Family - Alphaproteobacteria  stacked graphs of taxa composition}
# melt to long format (for ggploting) 
# prune out phyla below 1% in each sample
wrc_family <- meta_merge %>%
  tax_glom(taxrank = "Family") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, Class == "Alphaproteobacteria", source!="bulk_soil") %>%                      arrange(Family)                                     # Sort data frame alphabetically by phylum

# wrc_family
# Set colors for plotting
family_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_family$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_family, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_family, aes(x = Family, y = Abundance, fill = i)) + 
  geom_bar(stat = "summary", fun.y = "mean") +
  facet_wrap(~treatment)+
  scale_fill_manual(values = family_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14))+
  ylab("Relative Abundance > 1%") +
  xlab("Family Name")+
  labs(fill="Treatment") +
  ggtitle("Alphaproteobacteria Composition of \n Bacterial Communities") 

wrc_plot + coord_flip() 

p <- aov(Abundance ~ source*treatment*ditch + (1|block), wrc_family)
plot(p)
anova(p)

wrc_family_means <- wrc_family %>%
  group_by(source, Family) %>%
  summarise_at(c("Abundance"), mean, na.rm=TRUE)

write.csv(wrc_family_means, file="../figures/figs_2018_WRC_SoilTeaDecomp/alpha_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/alpha_family_stack.pdf", width=15, height=9, dpi=300)
```

```{r phyloseq Family - Deltaproteobacteria stacked graphs of taxa composition}
# melt to long format (for ggploting) 
# prune out phyla below 1% in each sample
wrc_family <- meta_merge %>%
  tax_glom(taxrank = "Family") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, Class == "Deltaproteobacteria", source!="bulk_soil") %>%                      arrange(Family)                                     # Sort data frame alphabetically by phylum

# wrc_family
# Set colors for plotting
family_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_family$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_family, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_family, aes(x = Family, y = Abundance, fill = i)) + 
  geom_bar(stat = "summary", fun.y = "mean") +
  facet_wrap(~treatment)+
  scale_fill_manual(values = family_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank())+ 
  theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14))+
  ylab("Relative Abundance > 1%") +
  xlab("Family Name")+
  ggtitle("Deltaproteobacteria Composition of \n Bacterial Communities") 

wrc_plot + coord_flip() 

p <- aov(Abundance ~ source*treatment*ditch + (1|block), wrc_family)
plot(p)
anova(p)

wrc_family_means <- wrc_family %>%
  group_by(source, Family) %>%
  summarise_at(c("Abundance"), mean, na.rm=TRUE)

write.csv(wrc_family_means, file="../figures/figs_2018_WRC_SoilTeaDecomp/delta_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/delta_family_stack.pdf", width=15, height=9, dpi=300)
```

```{r phyloseq Family - Gammaproteobacteria  stacked graphs of taxa composition}
# prune out phyla below 1% in each sample
wrc_family <- meta_merge %>%
  tax_glom(taxrank = "Family") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, Class == "Gammaproteobacteria", source!="bulk_soil") %>%  # Filter out low abundance taxa
  arrange(Family)                                     # Sort data frame alphabetically by phylum

# Set colors for plotting
family_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_family$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_family, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_family, aes(x = Family, y = Abundance, fill = i)) + 
  facet_grid(~treatment)+
  geom_bar(stat = "summary", fun.y = "mean") +
  scale_fill_manual(values = family_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14))+
  ylab("Relative Abundance > 1%") +
  xlab("Family Name")+
  ggtitle("Gammaproteobacteria Composition of \n Bacterial Communities") 

wrc_plot + coord_flip() 

write.csv(wrc_family, file="../figures/figs_2018_WRC_SoilTeaDecomp/gamma_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/gamma_family_stack.pdf", width=15, height=9, dpi=300)
```


```{r phyloseq Family - Betaproteobacteria  stacked graphs of taxa composition}
# prune out phyla below 1% in each sample
wrc_family <- meta_merge %>%
  tax_glom(taxrank = "Family") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, Class == "Betaproteobacteria", source!="bulk_soil") %>% # Filter out low abundance taxa
  arrange(Family)                                     # Sort data frame alphabetically by phylum

# Set colors for plotting
family_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_family$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_family, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_family, aes(x = Family, y = Abundance, fill = i)) + 
  facet_grid(~treatment)+
  geom_bar(stat = "summary", fun.y = "mean") +
  scale_fill_manual(values = family_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14))+
  ylab("Relative Abundance > 1%") +
  xlab("Family Name")+
  ggtitle("Betaproteobacteria Composition of \n Bacterial Communities") 

wrc_plot + coord_flip() 

write.csv(wrc_family, file="../figures/figs_2018_WRC_SoilTeaDecomp/beta_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/beta_family_stack.pdf", width=15, height=9, dpi=300)
```

```{r phyloseq Family - Actinobacteria  stacked graphs of taxa composition}
# prune out phyla below 1% in each sample
wrc_family <- meta_merge %>%
  tax_glom(taxrank = "Family") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, Class == "Actinobacteria", source!="bulk_soil") %>% # Filter out low abundance taxa
  arrange(Family)                                     # Sort data frame alphabetically by phylum

# Set colors for plotting
family_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_family$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_family, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_family, aes(x = Family, y = Abundance, fill = i)) + 
  facet_grid(~treatment)+
  geom_bar(stat = "summary", fun.y = "mean") +
  scale_fill_manual(values = family_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14))+
  ylab("Relative Abundance > 1%") +
  xlab("Family Name")+
  ggtitle("Actinobacteria Composition of \n Bacterial Communities") 

wrc_plot + coord_flip() 

write.csv(wrc_family, file="../figures/figs_2018_WRC_SoilTeaDecomp/actino_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/actino_family_stack.pdf", width=15, height=9, dpi=300)
```

```{r phyloseq Family - Rhodospirillaceae genus stacked graphs of taxa composition}
# prune out phyla below 1% in each sample
wrc_family <- meta_merge %>%
  tax_glom(taxrank = "Genus") %>%                     # agglomerate at phylum level
  transform_sample_counts(function(x) {x/sum(x)} ) %>% # Transform to rel. abundance
  psmelt() %>%                                         # Melt to long format
  filter(Abundance > 0.01, Family == "Rhodospirillaceae", source!="bulk_soil") %>% # Filter out low abundance taxa
  arrange(Genus)                                     # Sort data frame alphabetically by phylum

# Set colors for plotting
family_colors <- c("darkgreen", "darkred","blue", "orange")
levels(wrc_family$source) <- c("Bulk soil", "Green Tea", "Rooibos Tea")
labels <- c("0" = "Wet", "1" = "Dry")

i <- with(wrc_family, interaction(source, ditch))

levels(i) <- c("Bulk soil - Wet", "Green tea - Wet", "Rooibos tea - Wet", "Bulk soil - Dry", "Green tea - Dry", "Rooibos tea - Dry")

# Plot 
wrc_plot<- ggplot(wrc_family, aes(x = Genus, y = Abundance, fill = i)) + 
  facet_grid(~treatment)+
  geom_bar(stat = "summary", fun.y = "mean") +
  scale_fill_manual(values = family_colors) +
  theme_bw()  +
  theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank()) + 
  theme(axis.title=element_text(vjust=1,size=14,face="bold"),
          axis.text=element_text(size=14), axis.text.x = element_text(size=14))+
  ylab("Relative Abundance > 1%") +
  xlab("Genus Name")+
  ggtitle("Rhodospirillaceae Composition of \n Bacterial Communities") 

wrc_plot + coord_flip() 

write.csv(wrc_family, file="../figures/figs_2018_WRC_SoilTeaDecomp/rhodo_composition.csv", row.names=TRUE)

ggsave(file="../figures/figs_2018_WRC_SoilTeaDecomp/rhodo_family_stack.pdf", width=15, height=9, dpi=300)
```

```{r soil properties}
#Read in design file
meta.env <- design
meta.env$ditch <- as.factor(meta.env$ditch)

#bulk soils, M, MF
by_treatment <- meta.env %>%
  filter(source == "bulk_soil") %>%
  group_by(treatment) %>%
  summarise_at(c("temp_degC", "moisture", "pH", "c_percent", "n_percent", "C_N", 
                 "n_nh4_ug_gsoil", "n_no3_ug_gsoil", "p_po4_hcl_ug_gsoil", 
                 "p_po4_w_ug_gsoil"), funs(mean,sd), na.rm = TRUE)
write.csv(t(by_treatment), file = "../data/soil_properties_trt.csv",row.names=TRUE)

by_treatment_ditch <- meta.env %>%
  filter(source == "bulk_soil") %>%
  group_by(treatment, ditch) %>%
  summarise_at(c("temp_degC", "moisture", "pH", "c_percent", "n_percent", "C_N", 
                 "n_nh4_ug_gsoil", "n_no3_ug_gsoil", "p_po4_hcl_ug_gsoil", 
                 "p_po4_w_ug_gsoil"), funs(mean,sd), na.rm = TRUE)
write.csv(t(by_treatment_ditch), file = "../data/soil_properties_trt_ditch.csv",row.names=TRUE)

library("lmerTest")
library("pbkrtest")

temp.lm <- lmer(temp_degC ~ treatment*ditch + (1|block), data = meta.env)
plot(temp.lm)
anova(temp.lm, type="II", ddf="Kenward-Roger")
#confint(temp.lm)

nh4.lm <- lmer(n_nh4_ug_gsoil ~ treatment*ditch + (1|block), data = meta.env)
plot(nh4.lm)
anova(nh4.lm, type="II", ddf="Kenward-Roger")
#confint(nh4.lm)

no3.lm <- lmer(n_no3_ug_gsoil ~ treatment*ditch + (1|block), data = meta.env)
plot(no3.lm)
anova(no3.lm, type="II", ddf="Kenward-Roger")
#confint(no3.lm)

pH.lm <- lmer(pH ~ treatment*ditch + (1|block), data = meta.env)
plot(pH.lm)
anova(pH.lm, type="II", ddf="Kenward-Roger")
#confint(pH.lm)

soilC.lm <- lmer(c_percent ~ treatment*ditch + (1|block), data = meta.env)
plot(soilC.lm)
anova(soilC.lm, type="II", ddf="Kenward-Roger")
#confint(soilC.lm)

soilN.lm <- lmer(n_percent ~ treatment*ditch + (1|block), data = meta.env)
plot(soilN.lm)
anova(soilN.lm, type="II", ddf="Kenward-Roger")
#confint(soilN.lm)

CN.lm <- lmer(C_N ~ treatment*ditch + (1|block), data = meta.env)
plot(CN.lm)
anova(CN.lm, type="II", ddf="Kenward-Roger")
#confint(CN.lm)

p_hcl.lm <- lmer(p_po4_hcl_ug_gsoil ~ treatment*ditch + (1|block), data = meta.env)
plot(p_hcl.lm)
anova(p_hcl.lm, type="II", ddf="Kenward-Roger")
#confint(p_hcl.lm)

p_h2o.lm <- lmer(p_po4_w_ug_gsoil ~ treatment*ditch + (1|block), data = meta.env)
plot(p_h2o.lm)
anova(p_h2o.lm, type="II", ddf="Kenward-Roger")
#confint(p_h2o.lm)

moisture.lm <- lmer(moisture ~ treatment*ditch + (1|block), data = meta.env)
plot(moisture.lm)
anova(moisture.lm, type="II", ddf="Kenward-Roger")
#confint(moisture.lm)


nh4 <- aov(meta.env$n_nh4_ug_gsoil ~ treatment, meta.env)
anova(nh4)

no3 <- aov(meta.env$n_no3_ug_gsoil ~ treatment, meta.env)
anova(no3)

ph <- aov(meta.env$pH ~ treatment, meta.env)
anova(ph)

temp_degC <- aov(meta.env$temp_degC ~ treatment, meta.env)
anova(temp_degC)

cpercent <- aov(meta.env$c_percent ~ treatment, meta.env)
anova(cpercent)

npercent <- aov(meta.env$n_percent ~ treatment, meta.env)
anova(npercent)

CN <- aov(meta.env$C_N ~ treatment, meta.env)
anova(CN)

p_hcl <- aov(meta.env$p_po4_hcl_ug_gsoil ~ treatment, meta.env)
anova(p_hcl)

p_h2o <- aov(meta.env$p_po4_w_ug_gsoil ~ treatment, meta.env)
anova(p_h2o)

moist <- aov(meta.env$moisture ~ treatment, meta.env)
anova(moist)
```